### 2022-10-04 manuscript/manuscript.tex
more intro

- Disentangling evolutionary processes from the patterns of genetic variation
across the genome. The structure of what can be under selection shapes these
patterns when selection acts; when drift acts, these patterns should not exist.

@@

>>>>

\section*{Introduction}

As genomes are transmitted through the generations, lifetime events selectively
alter what is passed to the next generation. While selection directly acts on
the genetic variants that cause heritable fitness differences among
individuals, neighboring regions experience an indirect contraction in genetic
variability. This byproduct is known as \emph{linked selection}, and it occurs
when selection changes a site's frequency on timescale faster than
recombination can decouple that site from its genetic background. Because
linked selection modifies the genetic variability in a regionally-correlated
way, it creates observable patterns in genomic data that are used to understand
the type of selection processes acting on genomes and estimate key evolutionary
quantities.

% , with characteristic length
% $\nicefrac{s}{r}$ basepairs, where $s$ is the heterozygous selection
% coefficient and $r$ is the per-basepair recombination rate. Sweeps were found
% in humans, \emph{Drosophila}, and other species (XXX), though it was unclear
% whether hitchhiking alone could explain the observed genome-wide correlation
% between recombination rate and genetic diversity in some species (XXX).

Much work has been done on developing mathematical theory for two linked
selection processes: hitchhiking and background selection. These are
distinguished by the evolutionary processes that generate fitness variation and
the resulting selection dynamics. Under the classic hitchhiking model, a new
beneficial mutation sweeps fixation, leading to a dip in genetic diversity on a
small spatial scale. Since the sweeps leave relatively narrow footprint, they
influence genome-wide diversity only when beneficial mutations are recurring
and recombination is relatively limited, as in \emph{Drosophila}. These models
assume sweeps are sufficiently rare as to not occur simultaneously, which is
consistent with the slow estimated rate of strongly beneficial mutations (XXX).
Overall, the classic hitchhiking model is simple because the underlying
selection process does not generate polygenic fitness variation, even though
this is type of variation plays a dominant role in adaptation (XXX).

By contrast, background selection is a consequence of the fitness variation
created by a steady influx of new deleterious mutations. The classic background
selection model (XXX) considers how the balance between mutation introducing
deleterious variants into the population and selection removing them, reduces
nearby genetic variability. A critical assumption of this model is that such an
equilibrium between mutation and selection exists, which is the case only when
these mutations have a large enough fitness cost to ensure that they cannot fix
in the population. Under this process, the number of deleterious mutations per
gamete reaches a Poisson stationary distribution with rate $\nicefrac{U}{s}$,
where $U = \mu_D L$ is the deleterious mutation rate ($\mu_D$ per basepair per
generation) in a region of $L$ basepairs for mutations with a heterozygous
selection coefficient of $s$ (XXX). The fraction of gametes without any
mutations is then $f_0 = \exp(-\nicefrac{\mu_D}{s})$. Without recombination,
any gamete with deleterious mutations is destined to bring its linked neutral
variation with it to extinction, effectively reducing diversity levels to those
expected in a neutrally-evolving population of size $f_0 N$.

Remarkably, predictions from simple background selection models with
recombination closely match large-scale patterns of human diversity and
divergence (XXX). Since background approximately rescales a neutral coalescent
process by $f_0$, the predicted diversity is reduced from neutral levels by a
factor $B = \nicefrac{N_e}{N}$, which is the ratio of effective population size
under background selection to the population size. In background selection
models with recombination, $B$ is determined by the number of conserved
basepairs and recombination in a region. Consequently the predicted reduction
along the genome (known as $B$ maps) mirror the regional density of conserved
elements under purifying selection and recombination rates. These $B$ maps have
been used in studies of XXX, are inputs into machine learning models that
predict functional constraint, and XXX. Furthermore, since new deleterious
mutations are perpetually flowing into populations $B$ maps may provide a
better null model for evolution than the standard neutral model.

While background selection models do predict diversity levels

% outline?
- Disentangling evolutionary processes from the patterns of genetic variation
across the genome. The structure of what can be under selection shapes these
patterns when selection acts; when drift acts, these patterns should not exist.

@@

% Understanding the extent to which linked selection influences other
% evolutionary processes is a central question in evolutionary genetics. Two
% central questions have emerged. First, while genetic drift is thought to be the
% main source of random fluctuations in neutral allele frequencies, linked
% selection can have a similar impact on neutral sites, increasing the variance
% in allele frequency change as if the population size were effectively smaller.
% Although linked selection causes significant reductions in genetic diversity in
% some species, it is unlikely to explain the narrow range of observed diversity
% across species \parencite{Buffalo2021-uf}. Second, by perturbing the
% frequencies of neighboring alleles, selection can interfere with their
% selection dynamics, known as Hill-Robertson interference (HRI, XXX). When
% recombination is limited, simultaneous selection across the genome reduces the
% odds beneficial mutations are fixed (HRI), effectively like selection was
% occurring in an smaller population size. The Hill-Robertson effect is thought
% to be weak in organisms in organisms with large recombination maps, perhaps
% because of selection for increased recombination.

% In organisms with limited recombination such as
% \emph{Drosophila}, estimates suggest background selection plays a marginally
% larger role than hard sweeps in reducing diversity \parencite{Elyashiv2016-vt}.

% closely predict large-scale patterns of
% diversity across the genome

% Since far more new mutations are deleterious than beneficial in functional
% regions, it would seem like background selection would have a dominant role in
% shaping genetic diversity. However, deleterious mutations segregate on a
% timescale inversely proportional to their selection cost (XXX). Strongly
% deleterious mutations eliminate neighboring genetic variability on the few
% chromosomes they may recombine onto before the eventual extinction, and
% consequently have very weak effects on larger spatial scales. Moderately weakly
% selected mutations can persist for much longer periods, allowing for more
% recombination before their eventual loss and produce reductions in diversity on
% shorter spatial scales.  The reduction in diversity under background selection
% is determined approximately by $\nicefrac{U}{R}$, where $U = \mu_D L$ is the
% number of deleterious mutations (at rate $\mu_D$ per basepair, per generation)
% in a conserved region of $L$ basepairs per generation, and $R = r L$, which is
% the total recombination per generation in Morgans. Since the deleterious
% mutation rate is relatively constant across conserved regions, the reduction
% due to background selection closely matches the composition of conserved
% regions and recombination along the genome. Remarkably, simple mathematical
% models of background selection closely predicts levels of human-chimp
% divergence at the megabase scale (XXX) from genome-wide patterns of
% recombination and regions conserved between species (McVicker).

% This leads to some of the unidentifiability of hitchhiking versus background
% selection.

% and the reduction is determined
% by

% While the rate
% new beneficial mutations enter the population is low (XXX),

% The spatial scale of background selection is

% kkkkkk

% not all
% adaptation occurs from such new beneficial mutations sweeping to fixation (XXX)

% Other processes also act
% on beneficial fitness variation to engender linked selection, such as polygenic
% selection after an optimum shift \parencite{Buffalo2019-qs,Buffalo2020-my} or
% soft sweeps from standing variation (XXX).

% In turn, statistical methods have use this theory for a variety of wide-ranging
% empirical problems.  Early on, researchers

% While it is difficult to differentiate the selective processes
% leading to linked selection, studies indicate background selection

% hitchhiking events where a low-frequency

% studies
% in \emph{Drosophila} and humans indicate background selection has a large role
% in shaping patterns of diversity, especially at large spatial scales (i.e.
% megabases).

% Another central empirical goal is to estimate the relative strengths of the
% these two selection processes on levels of genetic diversity (Eyalshiv). Since
% the effects of linked selection are determined in part by mutation rate and
% distribution of fitness effects (DFE), these statistical methods also estimate
% these important evolutionary parameters from genomic data. These estimates
% should, in theory, align with others using different approaches, such as
% trio-based estimates of the mutation rate and estimates of the DFE from
% site-frequency spectra.

% Our focus here is to begin to fix some deficiencies in the background selection
% model. In particular,
